### 2021-11-29 rCode/cd_test.Rmd
CD adding maps and other visuals

---
output:
  pdf_document: default
  html_document: default
---
```{r,  echo=F, message = F, results = "asis"}

library(tidyverse)
# show map
model_input <- readRDS("../model_3_input.rds")
model_results <- readRDS("../model_3_results.rds")

# load voting data
library(data.table)
voting <- fread("C:\\Users\\darin\\OneDrive\\Documents\\school\\MPP1 Fall Semester\\GOV 2001\\Replication File\\cd\\voting data.csv")

model_input <- model_input %>%
  filter(is.finite(bartik_leo5_w2),
         is.finite(mfgNetChange_white),
         is.finite(mfgNetChange_nonwhite))

p75 <- quantile(model_input$mfgNetChange_white, (.75))

model_input$P_ddem_1 <- model_results$fitted.values

model_input <- model_input %>%
  mutate(
         CF_netchange = p75,
         CF_d_netchange = CF_netchange - mfgNetChange_white,
         CF_ddem_change = CF_d_netchange*model_results$coefficients[1],
         P_ddem_2 = P_ddem_1 + CF_ddem_change) %>%
  dplyr::select(id, state_fips, state, state_name, county_fips,
                county, year, dem_votes_pct1, ddem_votes_pct1,
                P_ddem_1, mfgNetChange_white, mfgNetChange_total,
                mfgNetChange_nonwhite,
                CF_netchange,
                CF_d_netchange,
                CF_ddem_change,
                P_ddem_2)

voting <- voting %>%
  dplyr::select(year, state, county_name, county_fips,
         party, candidatevotes) %>%
  filter(party == "DEMOCRAT" | party == "REPUBLICAN") %>%
  group_by(year, county_fips) %>%
  summarize(total_dem_rep_votes = sum(candidatevotes),
         dem_votes = sum(candidatevotes[party == "DEMOCRAT"]),
         rep_votes = sum(candidatevotes[party == "REPUBLICAN"])) %>%
  mutate(dem_vote_share = dem_votes/total_dem_rep_votes)

model_input <- model_input %>% mutate(
  c_fips_long = case_when(
    str_length(as.character(county_fips)) == 1 ~
      paste0("00", as.character(county_fips)),
    str_length(as.character(county_fips)) == 2 ~
      paste0("0", as.character(county_fips)),
    TRUE ~ as.character(county_fips)),

  s_fips_long  = case_when(
    str_length(as.character(state_fips)) == 1 ~
      paste0("0", as.character(state_fips)),
    TRUE ~ as.character(state_fips)),

  fips_long = paste0(s_fips_long, c_fips_long),
  fips_numeric = as.numeric(fips_long))

model_input <- left_join(model_input, voting,
                          by = c("fips_numeric" = "county_fips",
                                 "year" = "year"))

model_input <- model_input %>%
  mutate(CF_dem_votes = dem_votes +
           CF_ddem_change*total_dem_rep_votes,
         CF_rep_votes = rep_votes - CF_ddem_change*total_dem_rep_votes)

order = c("-2% or greater", "-1% to -2%", "-0.5% to -1%",
          "-0.25% to -0.5%", "0% to -0.25%", "0% to +0.25%")

state_summary <- model_input %>%
  group_by(state_name, state) %>%
  summarize(actual_trump_votes = sum(rep_votes),
            actual_dem_votes = sum(dem_votes),
            CF_trump_votes = sum(CF_rep_votes),
            CF_dem_votes = sum(CF_dem_votes)) %>%

  mutate(changed_outcome  =
           ifelse(
             CF_dem_votes > CF_trump_votes &
               actual_trump_votes > actual_dem_votes,
             1, 0),
    actual_trump_vs =
      actual_trump_votes/(actual_trump_votes+ actual_dem_votes),
    CF_trump_vs =
      CF_trump_votes/(CF_trump_votes+CF_dem_votes),
    delta_vs = CF_trump_vs - actual_trump_vs) %>%
  mutate(
    delta_vs = case_when(delta_vs < -.02 ~ "-2% or greater",
                         delta_vs < -.01 ~ "-1% to -2%",
                         delta_vs < -.005 ~ "-0.5% to -1%",
                         delta_vs < -.0025 ~ "-0.25% to -0.5%",
                         delta_vs < 0 ~ "0% to -0.25%",
                         delta_vs < .0025 ~ "0% to +0.25%",
                         delta_vs < .005 ~ "+0.25% to +0.5%",
                         delta_vs < .01 ~ "+0.5% to +1%",
                         TRUE ~ "+1% or greater")) %>%
  mutate(delta_vs_group = factor(delta_vs, levels = order))

colors =
   c("-2% or greater" = "#0059ff",
     "-1% to -2%" = "#4988fc",
     "-0.5% to -1%" = "#86b0fc",
     "-0.25% to -0.5%" = "#b8d1ff",
     "0% to -0.25%" = "#dee9fc",
     "0% to +0.25%" = "#ffeae8")

# map change in trump vote share
library(tigris)

# state shapefile
state_shape <- states(
  cb = TRUE,
  resolution = "20m"
) %>%
  filter(GEOID != "72") %>%
  shift_geometry()

mapping <- inner_join(state_shape, state_summary,
                      by = c("NAME" = "state_name"))

#devtools::install_github("yutannihilation/ggsflabel")
library(ggsflabel)

mapping_small <- mapping %>% filter(state != "DC")

mapping_small <- mapping_small %>%
  mutate(changed_outcome = ifelse(changed_outcome == 0,
                                  "No",
                                  "Yes"))

#  alt
flipped <- mapping_small %>% filter(changed_outcome == "Yes")
ggplot() +
  geom_sf(data = mapping_small, aes(fill = delta_vs_group), color = "grey", size = .01) +
  scale_fill_manual(values = colors) +
  labs(title = "Figure X:\nManufacturing Layoffs Led Trump to 2016 Victory",
       subtitle = "Change in Trump Vote Share Assuming Counterfactual Manufacturing Layoffs (25th Percentile)",
       fill = "Change in Trump Vote Share") +
  geom_sf(data = flipped, fill = NA, aes(color = changed_outcome), size = 1) +
  scale_color_manual(values = "#b32ab5", name = "Flipped Election") +
  labs(color = "Flipped Election") +
  coord_sf(datum = NA) +
  theme_minimal() +
  # geom_sf_text_repel(aes(label = state), data = mapping_small,
  #                    force = .00001,  seed = 10,
  #                    size = 2.5) +
  theme(  axis.title.x=element_blank(),
          axis.title.y=element_blank())

```