### 2021-11-30 rCode/Draft.Rmd
text changes

---
title: "Brother, can you spare a manufacturing job? White Voters reaction to Deindustrialization"
author: 'Catherine Darin and Zagreb Mukerjee'
date: '11/29/2021'
fontsize: 12pt
output:
  pdf_document: default
urlcolor: red
header-includes:
  - \usepackage{bm}
---

\renewcommand{\arraystretch}{1.1}
\renewcommand{\topfraction}{.85}
\renewcommand{\bottomfraction}{.7}
\renewcommand{\textfraction}{.15}
\renewcommand{\floatpagefraction}{.66}
\setcounter{topnumber}{3}
\setcounter{bottomnumber}{3}
\setcounter{totalnumber}{4}

## Introduction

What led to Donald Trump's surprising 2016 election victory? This paper examines the potential contribution of deindustrialization. Counties differ by exposure to manufacturing, which we use in an instrumental-variable approach to identify the effect of manufacturing job loss on change in Democratic vote share. We find that Democratic vote share falls in counties with high manufacturing job loss since 2004. Disaggregating this result by race shows that the democratic vote share fell where layoffs affected white populations, and rose where layoffs affected nonwhites. This suggests a racial component to how voters process economic hardship.

Base paper: Baccini and Weymouth 2021, "Gone For Good: Deindustrialization, White Voter Backlash, and US Presidential Voting." *APSR*.

## Outline

#### Framing story

To do: Qualitative description of the trajectory of XYZ county that swung from O to T and suffered big manufacturing layoffs. This will establish the stakes and connect this to the broader conversation.

### Substantive importance

Why did Trump win in 2016? A longstanding debate about relative importance of race and economic factors. Complicated by the interaction of the two. We hope to identify the importance of manufacturing job losses, both in aggregate and for different racial groups.

### Empirical strategy

Differential manufacturing exposure across counties allows for identification of the causal effect of deindustrialization on change in Dem vote share. Further differences in racial exposure to mfg layoffs allows for identification of the interaction between race and deindustrialization.

### Data and methods

Data is Census Quarterly Workforce Indicators, which break down county employment by industry, race and ethnicity. Augmented with public data on vote shares, employment, etc.

We look at total employment and manufacturing employment (NAICS 31-33, see extensions below). We then  compute net change in mfg jobs (long term job loss) from 2004 to 2015, relative to employment in 2004.

We might be worried about endogeneity for manufacturing job losses. So we use a shift-share instrument: we predict net job losses per worker with county-level manufacturing as a share of total employment, plus controls.

Controls: unemployment in 2015, college educated share of population, male share of population, layoffs in service sector, white share of population (to control for demographic trends). We want to know if there's something singular about manufacturing jobs.

### Extension Relative to Baccini and Weymoth (2021)

In their main model, Baccini and Weymouth looked at the effect on gross manufacturing job losses from 2012 to 2015 on the change in Democratic vote share in the 2016 presidential election.

Our key extension of Baccini and Weymouth's analysis is changing the quantity of interest from **gross** manufacturing job losses to **net** manufacturing job losses. For example, in Baccini and Weymouth's model, if a county lost 400 manufacturing jobs from 2012-2015 and also gained 400 manufacturing jobs over the same period, they would count 400 overall layoffs. Using our measure of *net* manufacturing job losses, we would consider this example a 0 net change in manufacturing employment. Given that job destruction and creation is a natural dynamic within the economy (i.e., creative destruction), we think that *net* job losses is a more meaningful measure.

Using *net* manufacturing job losses as our outcome, we replicate Baccini and Weymouth's main model, which uses 2012-2015 manufacturing (gross) job losses to explain the change in democratic vote share in 2016. Using their specifications, we find totally opposite results - that a *net* increase in white manufacturing jobs *increased* democratic vote share, while a *net* increase in non-white manufacturing jobs *decreased* democratic vote share. (See Table 3)

However, we do not trust the result of this specification, because, from 2012 to 2015, the national economy experienced a net increase in manufacturing jobs following the 2008 recession (See Figure 1). During this time of recovery, counties that experienced the most pronounced manufacturing job losses in prior years were also likely to experience the most pronounced recovery. Therefore, because manufacturing job losses pre-2012 are likely negatively correlated with manufacturing job gains from 2012-2015, it is misleading to segment the analysis data to just include job losses from 2012 to 2015. [To DO: Empirically show negative correlation describe here]

If we expand our measure of net manufacturing job losses to include the 2004 to 2011 period - which covers most of the period when the U.S. manufacturing economy was hit most intensely (See figure 1) -- we get a result that is consistent with Baccini and Weymouth's overall conclusion (See Table 4). That is, we find that net white manufacturing job losses decreased democratic vote share, while net non-white manufacturing job losses increased democratic vote share.

While our main takeaway is consistent with Baccini and Weymouth's story, it is important to highlight the substantive importance of the *timing* of manufacturing layoffs on political outcomes. Whereas Baccini and Weymouth's analysis suggests that economic changes lead to immediate political consequences, our analysis shows that economic changes can have lagged and variable effects on political outcomes.

[Note: to demonstrate our point about **variable** effects, In our next iteration of the analysis, we will show that manufacturing job losses had little explanatory effect on changes in democratic vote share in other elections (2012 and 2020).]

### Key Findings

We find that, in aggregate, manufacturing job losses were a driver of change in Democratic vote share in the 2016 election. Breaking it down by race, we see that

Moreover, we find that, given the evolution of the business cycle, considering only recent job losses can paint a misleading picture of the relationship between manufacturing job losses and changes in vote share. In fact, we find that net manufacturing job losses from 2004 to 2011 are crucial to explaining Trump's 2016 win. This finding suggests that economic changes, like manufacturing layoffs, can have lagged and variable political consequence.

### Potential Future Work

- Temporal: Finding a better way to characterize the potentially lagged effect of manufacturing job losses on Democratic votes. Incorporating study of the 2008, 2012, 2020 elections. 
- Disaggregation: Instead of instrumenting by manufacturing share, perhaps we can look at layoffs by industry in each county. While aggregate manufacturing employment may be endogenous, there's less reason to believe that specific industries would be (cf. Autor, Dorn and Hanson). This may also allow us to get further away from ecological inference problems.
- Trend-cycle estimation: There are techniques from econometrics and other places that could potentially be used to decompose gross layoffs into seasonal and non-seasonal components. 
- Mechanisms: survey work to examine where the racial difference in economic voting behavior comes from. Is it expressive? Is it because information is transmitted along racially segregated networks?

## Tables and Figures

```{r setup, echo=F, message = F, warning= F}
library(tidyverse)
library(stargazer)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(readxl)
library(scales)
library(tigris)

knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

### Manufacturing Layoffs Overtime

```{r}
# manufacturing layoffs overtime

national_ts <- read_xlsx("../National_Data_TS.xlsx")

national_ts <- national_ts %>%
  rename(year = 1,
         net_job_change = 2) %>%
  dplyr::select(year, net_job_change) %>%
  mutate(
    year = as.numeric(year),
    net_job_change = net_job_change*4) %>% group_by()

ggplot(data = national_ts %>% filter(year >= 1995), aes(x = year, y = net_job_change)) +
  geom_col() + scale_x_continuous(breaks=seq(1995,2019,2)) +
	scale_y_continuous(labels = label_number(suffix = "", scale = 1e-6)) +
  labs()

```

### Geographic Concentration of Layoffs

```{r,  echo=F, message = F, results = "asis"}

# load map
model_input <- readRDS("../model_5_input.rds")

# state shapefile
county_shape <- counties(
  cb = TRUE,
  resolution = "20m",
  progress_bar = F
) %>%
  shift_geometry() %>%
  mutate(STATEFP = as.numeric(STATEFP),
         COUNTYFP = as.numeric(COUNTYFP))

state_shape <- states(
  cb = TRUE,
  resolution = "20m",
  progress_bar = F
) %>%
  filter(GEOID != "72") %>%
  shift_geometry()

# group counties into sextiles 
model_input <- model_input %>%
  group_by() %>%
  mutate(ntile_mfg = as.character(ntile(mfgNetChange_total , 6)))

# find cutoffs

sum <- model_input %>%
  group_by(ntile_mfg) %>%
  summarize(min = min(mfgNetChange_total),
            max = max(mfgNetChange_total))

labels = c(
  "1" = "1% or Greater Gain in Mfg. Emp",
  "2" = "0% to 1% Gain in Mfg. Emp",
  "3" = "0% to 1% Loss in Mfg. Emp",
  "4" = "1% to 3% Loss in Mfg. Emp",
  "5" = "3% to 6% Loss in Mfg. Emp",
  "6" = "6% or Greater Loss in Mfg. Emp")

colors <- c("1" = "#778868",
            "2" = "#b5b991",
            "3" = "#f6edbd",
            "4" = "#edbb8a",
            "5" = "#de8a5a",
            "6" = "#ca562c")

#3d5941,#778868,#b5b991,#f6edbd,#edbb8a,#de8a5a,#ca562c
mapping_county <- inner_join(county_shape, model_input,
                      by = c("STATEFP" = "state_fips",
                             "COUNTYFP" = "county_fips"))

ggplot() +
  geom_sf(data = mapping_county, aes(fill = ntile_mfg), 
          color = NA, size = .01) +
  scale_fill_manual(values = colors,
                    labels = labels) +
  # add state lines
  geom_sf(data = state_shape, fill = NA, color = "black", size = .1) +
  coord_sf(datum = NA) +
  theme_minimal() +
  # geom_sf_text_repel(aes(label = state), data = mapping_small,
  #                    force = .00001,  seed = 10,
  #                    size = 2.5) +
  theme(  axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          text = element_text(size = 8)) + 
  labs(title = "Figure X:\nNet Change in Manufacturing Employment from 2004 to 2015",
       subtitle = "Normalized by 2004 Overall Employment Levels",
       fill = "Cumulative Change in Manufacturing\nEmployment (2004-2015) per\n2004 Worker")

```

### Descriptive Statistics
```{r,  echo=F}
descriptiveData <-  readRDS("../tableData.rds")

knitr::kable(descriptiveData$table04,
             caption = "Manufacturing Job Changes 2004-2015",
             booktabs = TRUE, digits = 2, format = 'latex')  %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))

knitr::kable(descriptiveData$tableOurs,
             caption = "Manufacturing Job Changes 2012-2015",
             booktabs = TRUE, digits = 2, format = 'latex')  %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))

```

## Figures

### Example

### Regression summaries

```{r,  echo=F, message = F, results = "asis"}

regResultList <- readRDS("../regresults.rds")

rownames(regResultList$secondStageModelNetCA$coefficients )[rownames(regResultList$secondStageModelNetCA$coefficients)=="firstStageModelNetCA$fitted.values"] <-  "Manufacturing Layoffs"
rownames(regResultList$secondStageModelNetC$coefficients )[rownames(regResultList$secondStageModelNetC$coefficients)=="firstStageModelNetC$fitted.values"] <-  "Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetCA$coefficients )[rownames(regResultList$secondStageModelWNetCA$coefficients)=="firstStageModelWNetCA$fitted.values"] <-  "White Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetC$coefficients )[rownames(regResultList$secondStageModelWNetC$coefficients)=="firstStageModelWNetC$fitted.values"] <-  "White Manufacturing Layoffs"

rownames(regResultList$secondStageModelWNetCA$coefficients 
)[rownames(regResultList$secondStageModelWNetCA$coefficients)=="mfgNetChange_nonwhite"] <-  "Nonwhite Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetC$coefficients
)[rownames(regResultList$secondStageModelWNetC$coefficients)=="mfgNetChange_nonwhite"] <-  "Nonwhite Manufacturing Layoffs"

rownames(regResultList$secondStageModelNetCA$beta )[rownames(regResultList$secondStageModelNetCA$beta) =="firstStageModelNetCA$fitted.values"] <-  "Manufacturing Layoffs"
rownames(regResultList$secondStageModelNetC$beta )[rownames(regResultList$secondStageModelNetC$beta)=="firstStageModelNetC$fitted.values"] <-  "Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetCA$beta )[rownames(regResultList$secondStageModelWNetCA$beta)=="firstStageModelWNetCA$fitted.values"] <-  "White Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetC$beta )[rownames(regResultList$secondStageModelWNetC$beta)=="firstStageModelWNetC$fitted.values"] <-  "White Manufacturing Layoffs"

rownames(regResultList$secondStageModelWNetCA$beta )[rownames(regResultList$secondStageModelWNetCA$beta)=="mfgNetChange_nonwhite"] <-  "Nonwhite Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetC$beta )[rownames(regResultList$secondStageModelWNetC$beta)=="mfgNetChange_nonwhite"] <-  "Nonwhite Manufacturing Layoffs"

names(regResultList$rseNetCA_2)[1] <- "Manufacturing Layoffs"
names(regResultList$rseNetC_2)[1] <- "Manufacturing Layoffs"
names(regResultList$rseWNetCA_2)[1] <- "White Manufacturing Layoffs"
names(regResultList$rseWNetC_2)[1] <- "White Manufacturing Layoffs"

names(regResultList$rseWNetCA_2)[names(regResultList$rseWNetCA_2) == "mfgNetChange_nonwhite"] <- "Nonwhite Manufacturing Layoffs"
names(regResultList$rseWNetC_2)[names(regResultList$rseWNetC_2) == "mfgNetChange_nonwhite"] <- "Nonwhite Manufacturing Layoffs"

stargazer(regResultList$secondStageModelNetCA,
          regResultList$secondStageModelNetC,
          regResultList$secondStageModelWNetCA,
          regResultList$secondStageModelWNetC,
          se = list(
            regResultList$rseNetCA_2,
            regResultList$rseNetC_2,
            regResultList$rseWNetCA_2,
            regResultList$rseWNetC_2
          ),
          title = "Effect of Manufacturing Layoffs on Democratic Vote Share",
          dep.var.labels = "Change in Share (2012-2016)",
          omit.stat = c("rsq","ser"),
          keep = c("Manufacturing Layoffs", "White Manufacturing Layoffs", "Nonwhite Manufacturing Layoffs"),
          add.lines = list(
            c("Controls For White Share/Service Layoffs", "No", "Yes", "No", "Yes")
          ),
          digits = 2,
          header = F,
          type = "latex"
)

```

```{r,  echo=F, message = F, results = "asis"}

rownames(regResultList$secondStageModelNetC04A$coefficients )[rownames(regResultList$secondStageModelNetC04A$coefficients)=="firstStageModelNetC04A$fitted.values"] <-  "Manufacturing Layoffs"
rownames(regResultList$secondStageModelNetC04$coefficients )[rownames(regResultList$secondStageModelNetC04$coefficients)=="firstStageModelNetC04$fitted.values"] <-  "Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetC04A$coefficients )[rownames(regResultList$secondStageModelWNetC04A$coefficients)=="firstStageModelWNetC04A$fitted.values"] <-  "White Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetC04$coefficients )[rownames(regResultList$secondStageModelWNetC04$coefficients)=="firstStageModelWNetC04$fitted.values"] <-  "White Manufacturing Layoffs"

rownames(regResultList$secondStageModelWNetC04A$coefficients
)[rownames(regResultList$secondStageModelWNetC04A$coefficients)=="mfgNetChange_nonwhite"] <-  "Nonwhite Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetC04$coefficients
)[rownames(regResultList$secondStageModelWNetC04$coefficients)=="mfgNetChange_nonwhite"] <-  "Nonwhite Manufacturing Layoffs"

rownames(regResultList$secondStageModelNetC04A$beta )[rownames(regResultList$secondStageModelNetC04A$beta) =="firstStageModelNetC04A$fitted.values"] <-  "Manufacturing Layoffs"
rownames(regResultList$secondStageModelNetC04$beta )[rownames(regResultList$secondStageModelNetC04$beta)=="firstStageModelNetC04$fitted.values"] <-  "Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetC04A$beta )[rownames(regResultList$secondStageModelWNetC04A$beta)=="firstStageModelWNetC04A$fitted.values"] <-  "White Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetC04$beta )[rownames(regResultList$secondStageModelWNetC04$beta)=="firstStageModelWNetC04$fitted.values"] <-  "White Manufacturing Layoffs"

rownames(regResultList$secondStageModelWNetC04A$beta )[rownames(regResultList$secondStageModelWNetC04A$beta)=="mfgNetChange_nonwhite"] <-  "Nonwhite Manufacturing Layoffs"
rownames(regResultList$secondStageModelWNetC04$beta )[rownames(regResultList$secondStageModelWNetC04$beta)=="mfgNetChange_nonwhite"] <-  "Nonwhite Manufacturing Layoffs"

names(regResultList$rseNetC04A_2)[1] <- "Manufacturing Layoffs"
names(regResultList$rseNetC04_2)[1] <- "Manufacturing Layoffs"
names(regResultList$rseWNetC04A_2)[1] <- "White Manufacturing Layoffs"
names(regResultList$rseWNetC04_2)[1] <- "White Manufacturing Layoffs"

names(regResultList$rseWNetC04A_2)[names(regResultList$rseWNetC04A_2) == "mfgNetChange_nonwhite"] <- "Nonwhite Manufacturing Layoffs"
names(regResultList$rseWNetC04_2)[names(regResultList$rseWNetC04_2) == "mfgNetChange_nonwhite"] <- "Nonwhite Manufacturing Layoffs"

stargazer(regResultList$secondStageModelNetC04A,
          regResultList$secondStageModelNetC04,
          regResultList$secondStageModelWNetC04A,
          regResultList$secondStageModelWNetC04,
          se = list(
            regResultList$rseNetC04A_2,
            regResultList$rseNetC04_2,
            regResultList$rseWNetC04A_2,
            regResultList$rseWNetC04_2
          ),
          title = "Effect of Manufacturing Layoffs on Democratic Vote Share (since 2004)",
          dep.var.labels = "Change in Share (2016-2012)",
          omit.stat = c("rsq","ser"),
          keep = c("Manufacturing Layoffs", "White Manufacturing Layoffs", "Nonwhite Manufacturing Layoffs"),
          add.lines = list(
            c("Controls For White Share/Service Layoffs", "No", "Yes", "No", "Yes")
          ),
          digits = 2,
          header = F,
          type = "latex"
)

```

### Counterfactual assessment of election
## (Using 2004-2015 Layoffs)

```{r,  echo=F, message = F, results = "asis"}

# load map
model_input <- readRDS("../model_5_input.rds")
model_results <- readRDS("../model_5_results.rds")

# load voting data
library(data.table)
voting <- fread("..\\voting data.csv")

# filter model input based on our regression filtering
model_input <- model_input %>%
  filter(is.finite(bartik_leo5_w2),
         is.finite(mfgNetChange_white),
         is.finite(mfgNetChange_nonwhite))

# find the 25th percentile in manufacturing layoffs (our measure) across counties
p25 <- quantile(model_input$mfgNetChange_white, (.25))

# find the predicted democratic vote share
model_input$P_ddem_1 <- model_results$fitted.values

model_input <- model_input %>%
  mutate(

    # calculate the counterfactual change in democratic vote share
    # by multiplying our regression coefficient against the difference
    # between observed manufacturing layoffs (our measure) and the
    # 25th percentile of layoffs (nationally)
         CF_netchange = p25,
         CF_d_netchange = CF_netchange - mfgNetChange_white,
         CF_ddem_change = CF_d_netchange*model_results$coefficients[1],
         P_ddem_2 = P_ddem_1 + CF_ddem_change) %>%
  dplyr::select(id, state_fips, state, state_name, county_fips,
                county, year, dem_votes_pct1, ddem_votes_pct1,
                P_ddem_1, mfgNetChange_white, mfgNetChange_total,
                mfgNetChange_nonwhite,
                CF_netchange,
                CF_d_netchange,
                CF_ddem_change,
                P_ddem_2)

# clear voting data
voting <- voting %>%
  dplyr::select(year, state, county_name, county_fips,
         party, candidatevotes) %>%
  filter(party == "DEMOCRAT" | party == "REPUBLICAN") %>%
  group_by(year, county_fips) %>%
  summarize(total_dem_rep_votes = sum(candidatevotes),
         dem_votes = sum(candidatevotes[party == "DEMOCRAT"]),
         rep_votes = sum(candidatevotes[party == "REPUBLICAN"])) %>%
  mutate(dem_vote_share = dem_votes/total_dem_rep_votes)

# prep model input data to be merged to voting data
model_input <- model_input %>% mutate(
  c_fips_long = case_when(
    str_length(as.character(county_fips)) == 1 ~
      paste0("00", as.character(county_fips)),
    str_length(as.character(county_fips)) == 2 ~
      paste0("0", as.character(county_fips)),
    TRUE ~ as.character(county_fips)),
  s_fips_long  = case_when(
    str_length(as.character(state_fips)) == 1 ~
      paste0("0", as.character(state_fips)),
    TRUE ~ as.character(state_fips)),

  fips_long = paste0(s_fips_long, c_fips_long),
  fips_numeric = as.numeric(fips_long))

# join model input to observed voting data
model_input <- left_join(model_input, voting,
                          by = c("fips_numeric" = "county_fips",
                                 "year" = "year"))

# calculate the counterfactual dem votes and rep votes
model_input <- model_input %>%
  mutate(CF_dem_votes = dem_votes +
           CF_ddem_change*total_dem_rep_votes,
         CF_rep_votes = rep_votes - CF_ddem_change*total_dem_rep_votes)

# create mapping categories
order = c(
  #"+2% or greater", "+1% to +2%",
          "+0.5% to +1%",
          "+0.25% to +0.5%", 
          "0% to +0.25%", 
          "0% to -0.25%")

state_summary <- model_input %>%
  group_by(state_name, state) %>%
  summarize(actual_trump_votes = sum(rep_votes),
            actual_dem_votes = sum(dem_votes),
            CF_trump_votes = sum(CF_rep_votes),
            CF_dem_votes = sum(CF_dem_votes)) %>%

  mutate(changed_outcome  =
           ifelse(
             CF_dem_votes > CF_trump_votes &
               actual_trump_votes > actual_dem_votes,
             1, 0),
    actual_trump_vs =
      actual_trump_votes/(actual_trump_votes+ actual_dem_votes),
    CF_trump_vs =
      CF_trump_votes/(CF_trump_votes+CF_dem_votes),
    delta_vs = CF_trump_vs - actual_trump_vs) %>%
  mutate(
    delta_vs = case_when(delta_vs < -.02 ~ "+2% or greater",
                         delta_vs < -.01 ~ "+1% to +2%",
                         delta_vs < -.005 ~ "+0.5% to +1%",
                         delta_vs < -.0025 ~ "+0.25% to +0.5%",
                         delta_vs < 0 ~ "0% to +0.25%",
                         delta_vs < .0025 ~ "0% to -0.25%",
                         delta_vs < .005 ~ "-0.25% to -0.5%",
                         delta_vs < .01 ~ "-0.5% to -1%",
                         TRUE ~ "-1% or greater")) %>%
  mutate(delta_vs_group = factor(delta_vs, levels = order))

colors =
   c(
     #"+2% or greater" = "#0059ff",
     #"+1% to +2%" = "#4988fc",
     "+0.5% to +1%" = "#86b0fc",
     "+0.25% to +0.5%" = "#b8d1ff",
     "0% to +0.25%" = "#dee9fc",
     "0% to -0.25%" = "#ffeae8")

# map change in trump vote share
library(tigris)

# state shapefile
state_shape <- states(
  cb = TRUE,
  resolution = "20m",
  progress_bar = F
) %>%
  filter(GEOID != "72") %>%
  shift_geometry()

mapping <- inner_join(state_shape, state_summary,
                      by = c("NAME" = "state_name"))

#devtools::install_github("yutannihilation/ggsflabel")
#library(ggsflabel)

mapping_small <- mapping %>% filter(state != "DC")

mapping_small <- mapping_small %>%
  mutate(changed_outcome = ifelse(changed_outcome == 0,
                                  "No",
                                  "Yes"))

#  alt
flipped <- mapping_small %>% filter(changed_outcome == "Yes")
ggplot() +
  geom_sf(data = mapping_small, aes(fill = delta_vs_group), color = "grey", size = .01) +
  scale_fill_manual(values = colors) +
  labs(title = "Figure X:\nWithout Manufacturing Layoffs, Democrats Would Have Won 2016 Election",
       subtitle = "Change in Democratic Vote Share Assuming Counterfactual Manufacturing Layoffs (25th Percentile)",
       fill = "Change in Democratic Vote Share") +
  geom_sf(data = flipped, fill = NA, aes(color = changed_outcome), size = 1) +
  scale_color_manual(values = "#b32ab5", name = "Flipped Election") +
  labs(color = "Flipped Election") +
  coord_sf(datum = NA) +
  theme_minimal() +
  # geom_sf_text_repel(aes(label = state), data = mapping_small,
  #                    force = .00001,  seed = 10,
  #                    size = 2.5) +
  theme(  axis.title.x=element_blank(),
          axis.title.y=element_blank())

```