### 2021-12-10 Final Draft/Final-Draft.tex
first draft of complete text

% AER-Article.tex for AEA last revised 22 June 2011
\documentclass[]{AEA}

% The mathtime package uses a Times font instead of Computer Modern.
% Uncomment the line below if you wish to use the mathtime package:
%\usepackage[cmbold]{mathtime}
% Note that miktex, by default, configures the mathtime package to use commercial fonts
% which you may not have. If you would like to use mathtime but you are seeing error
% messages about missing fonts (mtex.pfb, mtsy.pfb, or rmtmi.pfb) then please see
% the technical support document at http://www.aeaweb.org/templates/technical_support.pdf
% for instructions on fixing this problem.

% Note: you may use either harvard or natbib (but not both) to provide a wider
% variety of citation commands than latex supports natively. See below.

% Uncomment the next line to use the natbib package with bibtex
\usepackage{natbib}

% Uncomment the next line to use the harvard package with bibtex
%\usepackage[abbr]{harvard}

% This command determines the leading (vertical space between lines) in draft mode
% with 1.5 corresponding to "double" spacing.
\draftSpacing{1.5}

% For Pandoc highlighting needs

% Pandoc citation processing

\usepackage{hyperref}

\begin{document}

\title{Brother, Can You Spare a Manufacturing Job? How Voters React to
Deindustrialization}
\shortTitle{Brother, Can You Spare a Manufacturing Job?}
% \author{Author1 and Author2\thanks{Surname1: affiliation1, address1, email1.
% Surname2: affiliation2, address2, email2. Acknowledgements}}

\author{
  Catherine Darin\\
  Zagreb Mukerjee\thanks{
  Darin: Harvard University, \href{mailto:cdarin@hks.harvard.edu}{cdarin@hks.harvard.edu}.
  Mukerjee: Harvard University, \href{mailto:zagrebmukerjee@fas.harvard.edu}{zagrebmukerjee@fas.harvard.edu}.
}
}

\date{\today}
\pubMonth{12}
\pubYear{2021}
\pubVolume{}
\pubIssue{}
\JEL{}
\Keywords{}

\begin{abstract}
What led to Donald Trump's surprising 2016 election victory? This paper
examines the potential contribution of deindustrialization. We use an
instrumental-variable approach and a covariate-balancing approach to
identify the effect of manufacturing job loss on change in Democratic
vote share. We find that Democratic vote share falls in counties with
high manufacturing job loss since 2004. Disaggregating this result by
race shows that the Democratic vote share fell where layoffs affected
white populations, and rose where layoffs affected nonwhites. This
suggests a racial component to how voters process economic hardship.
\end{abstract}

\maketitle

This paper analyzes the relationship between deindustrialization, race,
and the election of Donald Trump in 2016. Previous analysis has
frequently focused on racial animus, trade, or the conjunction of the
two (\cite{Autor20}; \cite{che16}; \cite{BR21}). We extend the analysis
of \cite{Baccini21}, one of the first papers to examine the political
effects of deindustrialization while considering race and localization.
Deindustrialization - the transition of the U.S. economy away from
manufacturing - has led to widespread job loss and created adverse
knock-on effects for a wide swath of Americans. This experience has not
been uniform; rather, it has been concentrated in certain areas of the
country that were previously economically reliant on manufacturing
XXXXXXXXXXXX MAP FIGURE XXXXXXXXXXX. We believe the localized nature of
this analysis highlights the uneven and concentrated harms of
deindustrialization.

Along with \cite{Baccini21}, we find evidence that the experience of
deindustralization was associated with support for Donald Trump,
particularly among whites. However, we find that the related effects in
the 2016 election are not associated with recent manufacturing job loss;
rather, the opposite. Counties with manufacturing job losses from
2012-2015 tended to become more Democratic. We extend the analysis to
the 2004-2015 period, to capture the long-term effects of
deindustrialization on a region. Doing so, we find that regions
experiencing more manufacturing job losses swung towards Donald Trump in
2016, and that this occurred more when those losing jobs were white.

Our analysis extends \cite{Baccini21} in several ways. First, we change
focus from a measure of gross losses in manufacturing jobs to net
losses. As discussed in \ref{datasec}, we believe this to be a more
accurate measure of deindustrialization. Second, we change the temporal
focus. Rather than focusing on manufacturing job loss in 2012-2015, we
extend our analysis back to 2004. This is an important change. As shown
XXXXXXXXXXXXXX FIGURE XXXXXXXXXXXXXXX, the bulk of U.S. manufacturing
job loss took place prior to 2012; in fact, the 2012-2015 period was
characterized by a slight rebound in manufacturing jobs. During this
rebound, counties that experienced the most pronounced manufacturing job
losses in prior years were also likely to experience the most pronounced
recovery XXXXXXXXXXXXXXX FIGURE XXXXXXXXXXXXX. Because of this v-shaped
recovery, focusing only on 2012-2015 job losses/gains paints a distorted
picture. Finally, we augment the instrumental-variable approach with one
based on covariate balancing, which produces similar results.

\section{Data and Methods} 
\label{datamethods}

\subsection{Data} 
\label{datasec}

Following \cite{Baccini21}, our unit of analysis is the county. This
allows us to capture not only the direct effects of deindustrialization
on laid-off manufacturing workers, but also on the local economy. The
primary outcome variable is the county's change in Democratic vote share
between 2012 and 2016. We obtain this from XXXXXXXXX

We want to measure the causal relationship between deindustrialization
and the change in vote share. As a measure of deindustrialization, we
use the loss of manufacturing jobs as a share of the total
beginning-of-period employment in each county. Suppose at the beginning
of our period there were \(2000\) manufacturing workers in a county, and
\(8000\) non-manufacturing workers. At the end there are only \(1500\)
manufacturing workers. This is a loss of \(500/10000 = 5\%.\)

In this choice we diverge from \cite{Baccini21}, which uses
\textbf{gross} manufacturing job losses. For example, if a county lost
\(450\) manufacturing jobs from 2012-2015 and also gained \(400\)
manufacturing jobs over the same period, this would be \(450\)
\textbf{gross} job losses, but only \(50\) \textbf{net} job losses. The
\textbf{gross} measure captures several dynamics unrelated to
deindustrialization - for example, seasonal unemployment in a
food-manufacturing region, or workers moving between jobs. Thus we
believe the \textbf{net} job losses over a period are the more accurate
measure. It can be argued that the non-seasonal or non-structural
components of gross job losses are a useful covariate. Basic notions of
prospect theory suggest that the loss of a job and subsequent regaining
of another identical job may still create substantial discontent.
Constructing a better measure of gross job losses is an area for further
methodological and substantive research.

Data on job gains and losses are obtained, using an API, from the Census
Bureau's Quarterly Workforce Indicators (\cite{QWI}), which contains
(among other things) information about employment by industry and
county. These statistics are further disaggregated by race and
ethnicity. The Census Bureau obtains this data from a combination of
sources, such as administrative tax data and the U.S. Census.

XXXXXXXXXXXXXXXX sources of other data

\subsection{Methods} 
\label{methodssec}

We conduct two related tests of our hypothesis. Our first test involves
an instrumental-variable approach. A regression of change in Democratic
vote share on manufacturing job losses risks endogeneity, in case
counties with manufacturing job losses were otherwise predisposed to
turn towards Trump (a singular candidate, after all). To mitigate these
risks, \cite{Baccini21} uses a Bartik instrument, which we adapt (see
\cite{Bartik91}). This instrument essentially uses the cross-county
distribution of manufacturing employment as a source of exogenous
variation.

\[
\begin{aligned}
b_{j,c} &=& \frac{\text{Manufacturing Employment}_{j,c} \text{ at } t_0}{\text{Total Employment}_c \text{ at }t_0}  \\
&&  * \frac{\text{National Manufacturing Job Change}_{j,c} }{\text{Total National Employment at }t_0} \\
\end{aligned}
\]

Using this instrument, we then conduct a two stage regression. First we
estimate manufacturing job loss with the Bartik Instrument and a set of
county-level controls; then we use the estimated values of job losses to
predict Democratic vote share. In these regressions we control for
unemployment, service layoffs, the share of college-educated voters, and
the share of male voters. For some estimates, we also control for the
white share of the population. We also apply state fixed effects to
account for a wide variety of state-level changes from 2012-2016 (such
as voter suppression tactics). These controls are similar to those of
\cite{Baccini21}.

Our second approach involves attempting to balance the covariates of the
deindustrialization treatment. We use the same covariates as are used in
controls above, as well as manufacturing as share of the population and
total population. We apply the method of Covariate Balancing Propensity
Scores, as developed by \cite{Imai14}; this method generates weights to
reduce the correlation between covariates and treatment, while
addressing several of the issues with propensity score weights. Given
covariate-balancing weights, we then conduct a single-stage weighted
regression of change in Democratic vote share on manufacturing job loss,
using the same controls as above.

XXXXXXXXXXXXXXX BALANCE FIGURE??? XXXXXXXXXXXXXXXXXXXXXXXXXXX

\subsection{Results}

We find that counties experiencing deindustrialization from 2004-2015
are substantially more likely to vote for Trump. Further, we find that
this effect is largely attributable to whites. In other words, net white
manufacturing job losses decreased Democratic vote share, while net
non-white manufacturing job losses increased Democratic vote share. We
find the same effect when using covariate balancing, albeit of reduced
size.

The results can be found in XXXXXXXXXXXXX. The regression coefficient in
Model 4, for instance, means that a loss of \(10\) manufacturing jobs
per \(100\) workers in a county (a relatively extreme value) corresponds
to a loss of \(3.2 \%\) in Democratic vote share.

This coefficient is still somewhat difficult to interpret, and made
increasingly so by the winner-take-all system by which U.S. Electoral
College votes are allocated. As an alternative perspective, we examined
a range of counterfactual scenarios, in which counties experienced job
losses

XXXXXXXXXXXXXXXXXXXXXX EXPLAIN COUNTERFACTUAL

\subsubsection{Timing}

Restricting our analysis to only the 2012-2015 period produces an
opposite effect, in contrast to \cite{BW21} (when we use a net job loss
metric). The loss of Democratic vote share seems to be associated with a
gain in manufacturing jobs, both in general and among whites. We believe
this time-restricted analysis suffers from an omitted variable bias. As
discussed above, the 2012-2015 period can be broadly characterized as a
slight rebound from the manufacturing job losses in the 2008 financial
crisis. The counties that experience job losses from 2004-2012 are more
likely to have job gains in 2012-2015. An apparent relationship between
2012-2015 job gains and loss of Democratic vote share in fact represents
2004-2012 job losses that may lead to both the later job gains and the
loss of vote share. XXXXXXXXXXXXX FIGURE XXXXXXXXXXXX

This result suggests an important temporal dimension to the political
effects of deindustrialization - job losses in 2004-2012 make their
effects felt in 2016. In some ways, this result is more plausible than
if the 2012-2015 job losses were a significant covariate of voting. The
effects of job loss may take some time to be felt, at both individual
and community levels (see XXXXXXXXXXXXXX CITE). The mechanisms of this
relationship make a good topic for subsequent research.

\subsection{Conclusion}

As much research has noted, the national and global implications of
American deindustrialization are only beginning to be felt - from the
2016 election to isolationism to the opioid epidemic. Previous analyses
have often focused on choosing between explanations grounded in racial
issues or political-economic evolutions XXXXXXXXXXXXXXXXXXXXXxxx CITE.
We are joining a growing body of scholars searching for explanations at
the intersection of the two. We believe that deindustrialization is
fundamentally connected to particular places, and so the localized
approach adopted by \cite{BW21} offers a great deal of promise. Further
research might explore further ways in which geography has influenced
the rightward turn in the U.S., for example through the use of
techniques of spatial econometrics or the careful use of ethnography.

While our study is oriented towards history, deindustrialization is in
no way confined to the past. The trajectory of American manufacturing
employment is far from certain, with continued competition from abroad
and the perennial threat of unemployment. We believe the case of the
U.S. also offers insight that is applicable to the phenomenon of
`premature deindustrialization' in the developing world, and the
concurrent growth of populist parties in Hungary, Brazil, India and
elsewhere. XXXXXXXXXXXXXXX CITE XXXXXXxx

\begin{figure}

\caption{Caption for figure below.}
\begin{figurenotes}
Figure notes without optional leadin.
\end{figurenotes}
\begin{figurenotes}[Source]
Figure notes with optional leadin (Source, in this case).
\end{figurenotes}
\end{figure}

\nocite{stargazer}

\bibliographystyle{aea}
\bibliography{references}

\end{document}